---
title: "Spotify-top50-songs"
author: "Danni Chen - 705119383"
date: "6/4/2020"
output: html_document
---

```{r}
songs <- read.csv("../data/equally_sampled_songs.csv")
songs <- songs[, c(-1, -18)]
head(songs)
```


```{r}
table(songs$bad_influential)
```




### EDA stuff 
```{r}
library(ggplot2)
library(ggExtra)
```



Reference: http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization
```{r}
reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}
non_numerics <- c("genre", "track_name", "key", "mode", "bad_influential")
numerics <- !(colnames(songs) %in% non_numerics)

cormat <- round(cor(songs[, numerics]),2)
head(cormat)

# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
library(reshape2)
melted_cormat <- melt(cormat)
head(melted_cormat) 
# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
# Print the heatmap
print(ggheatmap)

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 2.5) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```






```{r}
ggplot(songs, aes(loudness, popularity)) + geom_point() + theme_bw()
```


#### plot year distribution

Reference https://cran.r-project.org/web/packages/ggExtra/readme/README.html

```{r}
piris <- ggplot(songs, aes(loudness, popularity, colour = genre)) +
  geom_point()
ggMarginal(piris, groupColour = TRUE, groupFill = TRUE)
```


```{r}
piris <- ggplot(songs, aes(loudness, popularity, colour = bad_influential)) +
  geom_point()
ggMarginal(piris, groupColour = TRUE, groupFill = TRUE)
```
#### plot a few highly correlated attributes
nrgy, acous, lyrics. lyrics.ease.score vs year val vs duration. year bmp vs db nrgy and db . dancinitu and dp , dancibility and violence


```{r}
colnames(songs)
```

## Why can genre predict popularity
#### plot genre. distribution colored by gender
```{r}
bp<-ggplot(songs, aes(x=genre, y=popularity, fill=genre)) +
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
bp
```


```{r}
bp<-ggplot(songs, aes(x=genre, y=popularity, fill=bad_influential)) +
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
bp
```


```{r}
bp<-ggplot(songs, aes(x=genre, y=loudness, fill=genre)) +
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
bp
```


```{r}
bp<-ggplot(songs, aes(x=genre, y=loudness, fill=bad_influential)) +
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
bp
```



```{r}
bp<-ggplot(songs, aes(x=genre, y=energy, fill=genre)) +
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
bp
```




```{r}
bp<-ggplot(songs, aes(x=genre, y=energy, fill=bad_influential)) +
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
bp
```

# There are pretty similar which is one major and minor is not a good predictor 

```{r}
ggplot(data = songs,
       mapping = aes(x = popularity, fill = genre)) +
  geom_histogram(alpha = 0.5, bins = 50) +
  labs(x = "Popularity Scores", y = "Count",
       title = "Popularity by genre") +
  facet_grid(. ~ bad_influential) +
  theme_bw()
```

```{r}
ggplot(data = songs,
       mapping = aes(x = loudness, fill = genre)) +
  geom_histogram(alpha = 0.5, bins = 50) +
  labs(x = "loudness Scores", y = "Count",
       title = "Popularity by genre") +
  facet_grid(. ~ bad_influential) +
  theme_bw()
```


```{r}
ggplot(data = songs,
       mapping = aes(x = energy, fill = genre)) +
  geom_histogram(alpha = 0.5, bins = 50) +
  labs(x = "Popularity Scores", y = "Count",
       title = "Popularity by genre") +
  facet_grid(. ~ bad_influential) +
  theme_bw()
```



```{r}
library(GGally)
ggpairs(data = songs,
      #  mapping = aes(color = genre),
        columns = c("acousticness", "loudness", "energy"),
        upper = list(continuous = wrap("cor", size = 2.5))
        )
```


